---
title: "TidyVerse Create_Data 607"
author: "Crystal Quezada"
date: "2024-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load and Clean the Data
The first step in our analysis is to load and clean the dataset using TidyVerse. For this, we'd analyze the age and party affiliation of Congress members. The main parties of interest are Democratic (100), Republican (200), and Independent (328). Party codes can be found at [VoteView.com](https://voteview.com/articles/data_help_parties).

```{r foundation}
library(tidyverse) 

# Load data
congressage_raw <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/congress-demographics/data_aging_congress.csv")   # read_csv 

# View raw data
head(congressage_raw)

# Clean data
congressage_clean <- congressage_raw %>%
  select(age_years, party_code) %>%     # select
  drop_na()        # drop_na()

# View clean data
head(congressage_clean)        # view first 6
```

# Visualize Age vs. Party Affiliation
Next, we create a scatter plot to visualize the relationship between age and party affiliation.

```{r visualization}
# Scatter plot of age vs party
ggplot(congressage_clean, aes(x = age_years, y = party_code)) +         # ggplot scatter plot
  geom_point() +
  labs(title = "Age vs Party Affiliation", x = "Age", y = "Party")
```

# Compute Summary Stat by Party
After visualizing the data, it's useful to compute the average age of all members in order to have a single measure. 

```{r summary}
# Mean stat by party
age_summary <- congressage_clean %>%            
  group_by(party_code) %>%          # group and summarize
  summarize(
    mean_age = mean(age_years),
  )

# View summary
age_summary
```

# Visualize Mean Age by Party
Finally, we create a scatter plot of the average age for each party to easily compare with other represented parties. 

```{r mean age plot}
# Scatter plot of mean age vs party
ggplot(age_summary, aes(x = party_code, y = mean_age)) +      # plot scatter plot with ggplot
  geom_point() + 
  labs(title = "Age vs Party Affiliation", x = "Party", y = "Mean Age")
```

# Conclusion

This vignette loads, cleans, and visualizes the data using TidyVerse to show the relationship between age and party affiliation. 
